How [preposition inanimate] becomes [\tsc{r}-pronoun postposition]\\ On \tit{waarmee} vs. \tit{met wat} in Dutch

### 2020-01-15 paper.tex
prefix vs suffix
kind of done, now still backtracking

@@

I assume that the \tit{'t} contains the ontological category \tsc{thing} \citep{kayne2005}. The feature Σ indicates that the pronoun is a weak pronoun. I leave possible number and gender features out because they do not play a role in this paper. The morpheme \tit{'t} can act as nominative, accusative and dative, as I showed in \ref{ex:tsubobj}. Taking this all together, \tit{'t} has the lexical entry given in \ref{ex:entryt}.

@@

This is due to the Elsewhere condition. The idea is that when two lexical entries are both candidates for spellout, the most specific is inserted.

@@

@@

@@

There is variation, between languages but also within languages.

@@

show lexical entry for this?

@@

@@

\ex.
\a. \begin{forest} boom
[\tsc{ins{}}
    [\tsc{f5}, roof]
]
{\draw (.east) node[right]{⇔ \tit{mit}}; }
\end{forest}\label{ex:mit}
\b. \begin{forest} boom
[\tsc{dat}P
    [\tsc{f}3]
    [\tsc{acc}P
        [\tsc{f}2]
        [\tsc{nom}P
            [\tsc{f}1]
        ]
    ]
]
{\draw (.east) node[right]{⇔ \tit{-em}}; }
\end{forest}\label{ex:em}

The distinction between prepositions and postpositions.

>>>>

 Pronouns in other genders alternate between nominative (non-oblique) and accusative/dative (oblique) in these contexts, illustrated in \ref{ex:hemsubobj}.

 \ex.\label{ex:hemsubobj}
 \ag. Hij staat in de hal.\\
  \tsc{3sg.m.nom} stands in the hallway\\
  `He is standing in the hallway.'
 \bg. Ik zie hem.\\
  I see \tsc{3sg.m.acc}\\
  `I see it.'
 \bg. Ik heb hem een klap gegeven.\\
  I have \tsc{3sg.m.acc} a hit given\\ with both alternatives. I work the proposal out with't  realizing all cases.

@@

I assume that the \tit{'t} contains the ontological category \tsc{thing} \citep{kayne2005}. The feature Σ indicates that the pronoun is a weak pronoun. I leave possible number and gender features out because they do not play a role in this paper. The morpheme \tit{'t} can act as nominative, accusative and dative, as I showed in \ref{ex:tsubobj}.\footnote{ Another possibility is to claim that \tit{'t} can only spell out \tsc{thing} and Σ and it combines with a zero suffix for the cases up to dative. This could be the same zero marker that full DPs combine with.

\ex.
\ag. De kast-∅ staat in de hal.\\
 the cabinet-\tsc{nom} stands in the hallway\\
 `The cabinet is standing in the hallway.'
\bg. Ik zie de kast-∅.\\
 I see the cabinet-\tsc{acc}\\
 `I see the cabinet.'
\bg. Ik heb de kast-∅ een klap gegeven.\\
 I have the cabinet-\tsc{dat} a hit given\\
 `I gave the cabinet a hit.'

The proposed account fares equally well with both alternatives. I work the proposal out with \tit{'t} realizing the cases up to the dative.} Taking this all together, \tit{'t} has the lexical entry given in \ref{ex:entryt}.

@@

This is due to the Elsewhere Condition. The idea is that when two lexical entries are both candidates for spellout, the most specific is inserted.

@@

Now we come to mee and met. the crucial distinction between these elements is that \tit{mee} is postposition and \tit{met} is a preposition. i show: (i) relation between these types of elements, (ii) how that is being modeled with movement, (iii) that post goes before pre (iv) how that is modeled in nano.

I shows 't spells out all features. this does not always happen. languages also use prepositions/prefixes and postpositions/suffixes to express case. and the relation between them is not arbitrary.

@@

can be modeled in the sense that a dp moves to a certain height.

@@

in german, dp moves above the dp. go back to dutch, hard to say above dative, because instrumental sometimes pre sometimes post.

There is variation, between languages but also within languages. another instance is bulgarian.

@@

@@

give lexical entry for that.
say why it does not work with full DPs and with animates.

so what happens with these instead? they get the prepostiion. prepositions have to be encoded differently in the syntax, because they ARE different, they occur before the expression. as you saw the -mee has a unary bottom, so it can only arise as a result of movement. pre elements have a binary bottom.

i'm going to crucially rely on the order in the spellout algoritm

MOVE TO LATER

@@
